Methodology for Assessing the State of the Practice for Domain X

### 2020-12-11 StateOfPractice/Methodology/Methodology.tex
Further work updating methodology document introduction

Research software uses computing to simulate mathematical models of real world
systems so that we can better understand and predict the system's behaviour.  A
small sample of some important examples of research software include the
following: designing new automotive parts, analysing the flow of blood in the
body, and determining the concentration of a pollutant released into the
groundwater.  As these examples illustrate, research software can be used for
tackling important problems that impact such areas as manufacturing, financial
planning, environmental policy, and the health, welfare and safety of
communities.

Given the importance of research software, scientists and engineers are pushing
for sustainable development practices.  This is evident from the existence of
such groups as SEI, BME.  Groups promoting sustainable practices are necessary
because the current state-of-the-practice for research software does not
incorporate state-of-the-art SE tools and methods
\citep{JohansonAndHasselbring2018}.  The lack of SE tools and methods
contributes to sustainability and reliability problems \citep{FaulkEtAl2009}.
Problems with the current state of the practice are evident from embarrassing
failures, like a retraction of derived molecular protein structures
\citep{Miller2006}, false reproduction of sonoluminescent fusion
\citep{PostAndVotta2005}, and fixing and then reintroducing the same error in a
large code base three times in 20 year \citep{MilewiczAndRaybourn2018}.  To
improve this situation, we need to more fully understand the current state of
the practice for research software.

To understand the current quality of SCS and to assist in developing methods for
measuring qualities, we will identify several domains within SCS and endeavour
to summarize the current state of the practice within those domains.  This
``state of the practice'' exercise will build off of prior work on
measuring/assessing the state of software development practice in several SCS
domains.  We will update the work that was done previously for domains such as
Geographic Information Systems \citep{SmithEtAl2018_arXivGIS}, Mesh Generators
\citep{SmithEtAl2016}, Seismology software \citep{SmithEtAl2018}, and
Statistical software for psychology \citep{SmithEtAl2018_StatSoft}.  We may
return to these domains and/or introduce new domains.  Potential
domains/collaborators are listed in the subsections below.

In the previous project, we measured 30 software projects for each domain, but
the measures were relatively shallow.  With the re-boot we will refine the llist
into a short list of the best projects for more detailed measurement.

>>>>

Research software uses computing to simulate mathematical models of real world
systems so that we can better understand and predict those system's behaviour.  A
small set of examples of important research software includes the
following: designing new automotive parts, analyzing the flow of blood in the
body, and determining the concentration of a pollutant released into the
groundwater.  As these examples illustrate, research software can be used for
tackling important problems that impact such areas as manufacturing, financial
planning, environmental policy, and the health, welfare and safety of
communities.

Given the importance of research software, scientists and engineers are pushing
for methods and tools to sustainably develop hiqh quality software.  This is
evident from the existence of such groups as the Software Sustainability
Institute (\href{https://www.software.ac.uk/} {SEI}) and Better Scientific
Software (\href{https://bssw.io/} {BSS}).  Groups such as these are necessary
for promoting sustainable practices because unfortunately the current
state-of-the-practice for research software often does not incorporate
state-of-the-art Software Engineering (SE) tools and methods
\citep{JohansonAndHasselbring2018}.  The lack of SE tools and methods
contributes to sustainability and reliability problems \citep{FaulkEtAl2009}.
Problems with the current state of the practice are evident from embarrassing
failures, like a retraction of derived molecular protein structures
\citep{Miller2006}, false reproduction of sonoluminescent fusion
\citep{PostAndVotta2005}, and fixing and then reintroducing the same error in a
large code base three times in 20 year \citep{MilewiczAndRaybourn2018}.  To
improve this situation, we need to first more fully understand the current state of
the practice for research software.

To understand the current quality of research software and to assist in
developing methods for measuring qualities, we will identify several domains
within research software and endeavour to summarize the current state of the
practice within those domains.  This ``state of the practice'' exercise will
build off of prior work on measuring/assessing the state of software development
practice in several research domains.  We will update the work that was done
previously for domains such as Geographic Information Systems
\citep{SmithEtAl2018_arXivGIS}, Mesh Generators \citep{SmithEtAl2016},
Seismology software \citep{SmithEtAl2018}, and Statistical software for
psychology \citep{SmithEtAl2018_StatSoft}.

In the previous project, we measured 30 software projects for each domain, but
the measures were relatively shallow.  With the re-boot we are refining the list
into a short list of the best projects for more detailed measurement.
